### 2024-07-23 Paper/paper2.tex
Updates from Overleaf

    \begin{abstract}
		{\small
        There is mixed evidence on whether not-for-profit hospitals differ meaningfully from for-profit (FP) hospitals, but what has not been considered is the effect leadership teams have on hospital behavior. In this paper, I investigate whether leadership team composition of NFP hospitals affects how similarly they behave to for-profits. I use a novel data set on NFP hospital executives in the US (2010-2014) to investigate whether the presence of clinically trained executives affects hospital behavior. In particular, I leverage pay-for-performance initiatives in the US enacted in 2012, and measure the  readmission and mortality rate responses between FP and NFP with and without clinically trained executives. I find that for-profits nad NFPs without clinical executives respond more aggressively to financial incentives on quality, consistent with theoretical predictions of firms that value profit relatively more than having high quality. 
		} 
	\end{abstract}

@@

  The objectives and behaviors of not-for-profit firms are challenging to determine, and an extensive literature has sought to understand them. Researchers have posited a variety of objective functions that could motivate NFP firms such as maximizing profit (for-profit in disguise), prestige, income, or a quality/quantity trade-off (\cite{steinberg1986revealed}). Empirical contributions to this literature center around investigating behaviors that differ between for-profit (FP), private NFP, and government firms (\cite{sloan2000not}). However, the literature has neglected to consider other relevant characteristics of hospitals that could affect objectives, such as the hospital's leadership team. In this paper, I investigate whether differences in observed behavior between for-profit and not-for-profit hospitals vary differentially by leadership team composition.

  Hospitals have been a common focus when seeking to understand NFP firms. This is because, first, there are different hospital ownership types within the industry, giving researchers the ability to compare the behaviors of NFPs and FPs. Second, private NFP hospitals make up 50\% of all hospitals in the US, and staff, on average, 207 beds. For comparison, FPs make up 36\% of hospitals and staff 107 beds (\cite{ASPE_2023}). Thus, NFP behavior in the hospital industry is relevant, and directly affects consumers of health care. However, it remains unclear what objective function drives NFP hospital behavior, and whether NFP hospitals are meaningfully different from FP hospitals (\cite{sloan2000not}; \cite{erus2002inferring}; \cite{deneffe2002not}; \cite{horwitz2009hospital}).

  While ownership is certainly relevant, there could be other characteristics correlated with hospital behavior independent of ownership type. Researchers have focused in particular on explaining variation in NFP behavior using variation in the market they exist in (\cite{duggan2002hospital}; \cite{horwitz2009hospital}). However, I argue that an observable and economically meaningful factor is the composition of top-level hospital managers. In large publicly traded firms, executive characteristics are shown to be correlated with firm performance and governance, indicating that there is more to firm behavior than ownership status (\cite{bertrand2003managing}; \cite{ahern2012changing}\cite{matsa2013female}).

  A policy relevant aspect of hospital behavior in the US has been the shift towards pay-for-performance incentives. While quality may or may not directly matter to hospitals, pay-for-performance incentives are such that quality of care directly impacts profit. For hospitals that already valued quality before pay-for-performance incentives, the change in incentives may not change quality much. However, if the hospital is more profit driven, the policy change should translate to a larger change in quality. Ownership status certainly plays a role in the valuations of profit vs. quality, but the question remains as to whether management also plays an important role. I contribute to our understanding of hospital and NFP behaviors by investigating whether occupational background of executives, specifically clinical training, affects hospital response to pay-for-performance incentives.

  Using publicly available tax forms, I construct a novel data set of NFP hospital executives from 2009-2015. I link hospitals in this data to characteristics in the American Hospital Association (AHA) survey and Hospital Compare data. I leverage various pay-for-performance initiatives in 2012 to investigate differential hospital responses between for-profits, and NFPs with and without without clinically trained executives.

  I estimate how the interaction between hospital type and pay-for-performance incentives affects readmission and mortality rates. Under the assumption that hospitals would have continued to behave similarly in the absence of the programs, and that leadership teams do not change endogenously with the programs, I identify the differential response to pay-for-performance incentives among hospital types. That is, I define a group of treatment and comparison hospitals, and estimate how outcomes differ between the two groups after the policy changes. I ultimately employ synthetic difference-in-differences to analyze a group of comparison hospitals with similar characteristics to treatment hospitals.\footnote{I also present results for the full sample and find qualitatively similar results.}

   While quality generally improves after pay-for-performance incentives for all hospital types, the most aggressive response is from for-profits and NFPs without clinical executives, who decrease readmission and mortality rates more so than NFPs with clinical executives. That is, the objectives of NFPs without clinically trained executives seem to align more with those of for-profits: placing more weight on profit than societal outcomes. Leveraging changes in executive teams, I decompose this effect into signalling (clinical executives signal underlying hospital preferences), or managing (clinical executives manage the hospital differently) effects. I find that the entire difference in response is driven by a managing effect.

    This paper contributes to three strands of literature. First, it adds to our understanding of NFP hospital objective functions. There is mixed evidence on whether NFP behavior differs from FP behavior in terms of costs, uncompensated care, technology adoption, and quality (\cite{sloan2000not}; \cite{eggleston2008hospital}; \cite{moscelli2018effect}; \cite{moscone2020public}). To this point, researchers have found that NFPs do not act as purely profit maximizers, but maximize some combination of profit and output (\cite{deneffe2002not}, \cite{chang2011nonprofit}). This paper adds to our understanding of inputs into not-for-profit hospital objective functions.

    Second, many researchers have documented a correlation between executive backgrounds, specifically for CEOs, and firm performance, starting with \citeauthor{bertrand2003managing} (\citeyear{bertrand2003managing}), who show that manager fixed effects are an important driver of many firm behaviors and decisions. Female board members are associated with better oversight and more women executives (\cite{matsa2011chipping}; \cite{adams2009women}), but are negatively correlated with firm value (\cite{ahern2012changing}). Having more female executives is correlated with female employee wages and corporate strategies (\cite{flabbi2019female}; \cite{matsa2013female}); young male CEOs tend to be more aggressive in mergers and acquisitions, while those with military experience are less aggressive (\cite{levi2010deal}; \cite{benmelech2015military}); CEOs with general ability tend to receive higher pay and perform better (\cite{kaplan2012ceo}; \cite{custodio2013generalists}; \cite{adams2018director}; \cite{frydman2019rising}). Further, Chief Diversity Officers are not found to have any effect on hiring more diversely in universities (\cite{bradley2022impact}).

    Due to data limitations, many of these studies focus on large, publicly traded firms in the US. \citeauthor{brickley2010board} (\citeyear{brickley2010board}) has the only study to my knowledge answering this question in the US NFP context. Using exogenous variation in expected Medicare profits, the authors find that having an internal board of directors increases CEO compensation, and having physician board members decreases public donations (\cite{brickley2010board}). Additionally, two studies focus on hospital performance in other countries. \citeauthor{janke2019impact} (\citeyear{janke2019impact}) uses data from England to study whether CEOs affect hospital production, and find no association (\cite{janke2019impact}). However, \citeauthor{otero2022managers} (\citeyear{otero2022managers}) investigates the role of CEOs in public hospitals and Chile, and finds an 8\% decrease in mortality rates for hospitals with top managers (\cite{otero2022managers}). I contribute to this literature by considering the executive team as a whole rather than just CEOs, by investigating a context which is not largely studied but is policy relevant, NFP US hospitals, and by leveraging a unique executive characteristic pertinent to that context.

    Finally, I contribute to our understanding of how providers respond to pay-for-performance incentives in health care. The most in depth study of how hospitals respond to HRRP is \citeauthor{gupta2021impacts} (\citeyear{gupta2021impacts}), who finds that hospitals decreased readmissions by 5\% and mortality rates by 2\% on average as a result of the program, confirming prior studies (\cite{mellor2017does}; \cite{ziedan2018essays}; \cite{ody2019decreases}; \cite{gupta2021impacts}). Around 40\% of the decrease in readmissions is due to selective patient practices. Research on HVBP generally finds that the program had no effect on underlying hospital quality (\cite{us2015hospital}; \cite{norton2018moneyball}; \cite{friedson2019so}). This paper contributes to our understanding of how characteristics of hospitals affect response to pay-for-performance incentives.

    \section{Setting}

    \subsection{Not-for-Profit Hospital Executives}

    Not-for-profit firms in the US differ from for-profits mainly in that for-profit firms distribute revenue to stakeholders, while not-for-profits reinvest profit back into the firm. Additionally, NFP firms are generally expected to set goals to better society due to government tax breaks received.

    Hospitals are governed by a board of directors, whose role is to set broad goals and strategies, and provide general oversight. The board selects executives, the highest level of management of a firm, to carry out day-to-day operations. An executive team usually consists of at least a Chief Executive Officer (CEO) and Chief Financial Officer (CFO), but there is variation in how firms organize executive teams. It is not uncommon for hospital executives to specialize in health care settings by earning a degree in health care management, or an MBA specific to health care. There are executive positions that are often filled by someone with clinical training, such as a Chief Medical Officer (CMO), and medical doctors pursue other top managerial positions as well, such as CEO or president. While some doctors earn additional degrees before stepping into an executive role, this is not a necessary condition to becoming a physician executive.

    Hiring physician executives can be controversial. On one hand, physicians bring a unique combination of clinical and administration expertise, and they have insights that can potentially improve hospital outcomes (\cite{Stajduhar_2023}, \cite{Ahmed_2022}). Alternatively, physicians may not always be adequately trained to enter into high level management positions, and therefore worsen hospital outcomes (\cite{HarvardBusinessReview2018}).

    \subsection{Pay-for-Performance Policies}\label{sec:hrrp}

@@

	\section{Data}\label{sec:data}

    Hospital executive teams are understudied partly due to the inaccessibility of granular data. I construct a novel data set of not-for-profit hospitals in the US that contains names, titles, and positions of all board members and executives tied to the hospital from 2009-2015. I gather this data from publicly available Tax Form 990s, which every sufficiently large not-for-profit files each year in the US. To my knowledge, this is the first large scale gathering of executive names from these forms.\footnote{\citeauthor{brickley2010board} (\citeyear{brickley2010board}) collects compensation data for a select number of hospitals.}

    The historical tax forms for all not-for-profits are housed by ProPublica.\footnote{https://projects.propublica.org/not-for-profits/} I use the NonProfit Explorer API to extract Employee Identification Numbers (EIN) of NFPs with a hospital National Taxonomy of Exempt Entities (NTEE) code. After filtering out associations and specialty hospitals, there are approximately 3,000 NFP EINs. Linking these EINs to other hospital information is crucial for the analysis. I therefore rely on matching by location and name of hospital in both the tax forms and AHA data. I confidently match 1200 EINs to an AHA identification number based on exact name matches within the same state. I assess the observable differences between matched and non-matched hospitals in Appendix \ref{app:matched}, and conclude that the samples are overall similar apart from an under-sampling of hospitals that belong to systems.

    For the purposes of this paper, I focus mainly on executive teams as a whole instead of specific positions. The main reason for this is because the organization of teams varies so drastically across hospitals that a specific position, even if it exists in two different hospitals, can be meaningfully different. I drop anyone who is solely a board member, and focus on those who have executive in their title, or are labeled as a president or vice president of a department. I create various hospital-level characterizations based on their executive team, such as the existence of a clinically trained executive, the number of clinically trained executives, the number of total executives, and whether the hospital has a Chief Medical Officer. Additionally, I examine changes in propensity to hire a clinical executive. Summary statistics for these variables are shown in Table \ref{leader_sumstats}. On average, NFP hospitals in my sample have 4 executives and .4 clinically trained executives. Sixty-one percent of the sample never changes propensity to hire a clinically trained executive, and 33\% of NFPs in the sample have a Chief Medical Officer at some point.

    I combine the resulting sample of not-for-profits with all for-profits in the AHA survey. The outcomes I consider are readmission and mortality rates, which are directly targeted in the pay-for-performance policies, and are found in the CMS Hospital Compare data. I combine rates for pneumonia, AMI, and heart failure (the relevant HRRP conditions) into a weighted average based on the number of patients in each condition. Additionally, I include several relevant hospital characteristics from the AHA survey: number of beds, and whether the hospital is an academic medical center, is physician owned, or is system affiliated. Case mix index, a measure of overall patient complexity, comes from the CMS Case Mix Index files. Whether the hospital was penalized under HRRP or received payments under HVBP is found in the HCRIS data.

    % In Table \ref{sumstats}, I present summary statistics for hospital variables. There are 3,766 total hospitals in the sample, with 150 beds on average. Hospitals in the sample are less likely to be penalized for AMI than either heart failure or pneumonia under the HRRP program. On average, 20\% of patients with one of these conditions would be readmitted. While readmission rates for the specific conditions are similar, the highest likelihood of readmission is for heart failure patients. Alternatively, AMI is the most dangerous condition in terms of mortality rates.

    %\import{Tables}{overall_sumstats.tex}

    In Table \ref{tab:sumstats_samples}, I present a table of means of relevant variables for several hospital samples used throughout the analysis. First, there are 708 for-profit hospitals in the sample, shown in the left-most column. Notably,

    There are several samples of hospitals that are relevant for the analysis. First, there are all for-profits and all not-for-profits. Further, I subset the sample of NFPs into whether they have a clinical executive at some point in the sample. Finally, I characterize a NFP leadership team as being stable or not. A stable team means the hospital did not change their propensity to hire a clinical executive over time. I present a table of means for each of these sub-samples in Table \ref{tab:sumstats_samples}. Approximately 850 NFPs have leadership information, with 384 of them hiring an MD executive at some point in the sample. I compare these NFPs with the 708 for-profits from the AHA Survey. For-profits and NFPs are roughly the same bed size, but NFPs see more patients in each of the three relevant conditions. For-profits are more likely to be penalized, but also more likely to receive HVBP incentives (potentially for improvement in quality). For-profits have higher readmission rates, and higher mortality rates in all categories except heart failure. There are also key differences in not-for-profits with and without clinically trained executives. Those with MDs are larger, more likely to be penalized, have higher uncompensated care, and more complex patients. This holds for stable leadership teams as well, though the number of NFPs with a clinical executive drops significantly. I discuss the implications of these differences for identification in Section \ref{sec:model}.

@@

    I plot weighted average readmission and mortality rates across time for different hospital types in Figure \ref{fig:weighted_read_mort_graph}, where I adjust both rates by baseline patient complexity. Pay-for-performance incentive changes are shown as a dotted line between 2011 and 2012. For-profit hospitals have the highest readmission rates prior to 2012. While the readmission rate trend behaves similarly across hospital types prior to the programs, there are notable trend differences after the programs. NFPs with clinically trained executives show a more mitigated decrease in readmissions compared to the other types. In mortality rates, NFPs with a clinically trained executive have the lowest rates before the programs, but differentially continue with an increasing trend in mortality after the programs. While this is purely descriptive evidence, the trends in outcomes point to more similar behavior between for-profits and NFPs without a clinically trained executive in comparison to NFPs with a clinically trained executive.

@@

    I model hospital behavior under two simplified worlds: one where quality does not directly affect profit, and one where it does. Pay-for-performance incentives essentially incorporate performance directly into a hospital's profit function. I directly model performance as quality of care, which is inversely related to readmission and mortality rates. While the literature speculates about the true form of NFP objective functions, we can generally think of hospitals choosing some linear combination of profit and societal benefit, without knowing how much weight the hospital places on each. Intuitively, a hospital that places only weight on profit (for-profit type) will only care about quality when it enters into the profit function, and therefore will respond more drastically to pay-for-performance than a hospital who already cared about quality through societal benefit before the incentives. Therefore, hospital responses to pay-for-performance incentives reveal information about their underlying preference on profit vs. societal outcomes.

    \noindent where $R$ is net revenue, $c_{\theta}(.)$ is an increasing cost function specific to quality $\theta$, $u(.)$ is extra utility gained from quality, which is increasing and concave in $\theta$, and $\alpha\in[0,1]$ captures the weight placed on profit vs. societal benefit. There is an implicit stay open condition throughout the model. Taking the first order condition yields

    \section{Estimating Differences in Hospital Response}

    \subsection{Method}

    In light of the theoretical predictions, I test empirically whether different hospital types respond to pay-for-performance incentives differently. I present here a two-way fixed effects specification, where the key variable is an interaction between a certain hospital type and time periods after program enactments. Ultimately, I employ synthetic difference-in-differences, which is estimating the same model with weights that more heavily factor similar trending control groups and time periods. However, for clarity, I present the unweighted specification:

    \begin{equation}
    \label{eq:forprofit}
    y_{ht} = \beta \text{ treat}_{ht} + \gamma_{h} + \delta_t + \epsilon_{ht},
    \end{equation}

    \noindent where $y_{ht}$ is one of the outcome variables discussed in Section \ref{sec:data}, and $\gamma_h$ and $\delta_t$ are hospital and time fixed effects, respectively. The independent variable of interest, treat$_{ht}$, captures the effect of being a certain hospital type after incentives changed in 2012. That is, I interact an indicator for hospital type with an indicator for year post 2012 for various hospital types. First, I define treat so that for-profit hospitals are being compared to subsets of not-for-profits (by limiting the sample to the relevant treatment and comparison group). Then, I limit to only not-for-profits and define treat as those without a clinical executive. Specifically, I define treat$_{ht}$ and limit the sample in four ways, shown in Figure \ref{fig:spec}.

    An important distinction is whether the classification of NFPs by leadership team is time-varying. For overall ownership status, switching is rare, and I limit to those who never switch. However, hospital executive teams are dynamic, and whether or not a hospital employs a clinically trained executive in each period may change. Therefore, I first present effects of having a clinically trained executive among stable executive teams, i.e., NFPs with no change in their propensity to hire a clinically trained executive.

\begin{figure}[ht!]
    \begin{center}
    \caption{\label{fig:spec}Treated and Comparison Groups}
 \begin{tabular}{| m{18em} |}
 \hline
 Combination 1:\\ [0.5ex]
 \hline\hline 
 \vspace{2mm}
 Treat Group: \hspace{23.5mm} For-Profit \\
 \vspace{2mm} 
 Comparison Group: \hspace{11.5mm} Stable NFPs  \\
 [1ex]
 \hline
 \end{tabular}
\hfil   %<---
 \begin{tabular}{|m{18em}|}
 \hline
 Combination 2:\\ [0.5ex]
 \hline\hline
 \vspace{2mm}
 Treat Group: \hspace{13.5mm} For-Profit \\
 \vspace{2mm}
 Comparison Group: \hspace{1.5mm} Stable NFP w/out MD  \\
 [1ex]
 \hline
 \end{tabular}

\medskip

\vspace{2mm}

  \begin{tabular}{|m{18em}|}
 \hline
 Combination 3:\\ [0.5ex]
 \hline\hline
 \vspace{2mm}
 Treat Group: \hspace{17mm} For-Profit \\
 \vspace{2mm}
 Comparison Group: \hspace{5mm} Stable NFPs w/ MD  \\
 [1ex]
 \hline
 \end{tabular}
\hfil   %<---
  \begin{tabular}{|m{18em}|}
 \hline
 Combination 4:\\ [0.5ex]
 \hline\hline
 \vspace{2mm}
 Treat Group:  \hspace{13mm} Stable NFPs w/out MD \\
 \vspace{2mm}
 Comparison Group:  \hspace{1mm} Stable NFPs w/ MD  \\
 [1ex]
 \hline
 \end{tabular}
 \end{center}
 \end{figure}

 \subsection{Identification}

 To identify the causal effect of hospital type on hospital program response, I rely on several assumptions. First, I assume that, absent the program enactments, treat and comparison group types would have behaved similarly in outcomes. While I see no evidence of pre-trends, Table \ref{tab:sumstats_samples} suggests that hospitals with a clinically trained executive are different than other hospitals along multiple characteristics. There may be concern that these differences between hospital types could be driving the results. Therefore, my main specification uses the synthetic difference-in-differences method established in \citeauthor{arkhangelsky2021synthetic} (\citeyear{arkhangelsky2021synthetic}). This method creates unit weights that more heavily factor comparison hospitals that are similar to the treat group hospitals prior to the programs. It also weighs time periods that balance pre-programs and post-programs for the comparison hospitals. Then, a simple two way fixed effects estimator is used with these weights, yielding a more robust average treatment effect that relies less on a strong parallel trends assumption (\cite{arkhangelsky2021synthetic}). I present the synthetic diff-in-diff estimates as the main results, and they are largely similar to the standard two way fixed effects results with no weights, shown in Appendix \ref{app:fullsample}. Further, the conceptual identification is the same as a standard difference-in-differences.

@@

 The estimated difference in readmission rate trends for all hospital types in response to the programs are shown in Figure \ref{fig:read_synth_plot}. Each panel corresponds to a particular definition of the treat group of hospitals and limitation of comparison group hospitals.

>>>>

    \begin{abstract}
		{\small
         Firm leaders are shown to be correlated with firm decisions and performance. In this paper, I investigate this topic in the context of hospital executive teams. Specifically, I empirically the effect of clinically trained executives on hospital behavior in response to an incentive shock. In particular, I leverage policies changing quality incentives for hospitals in the US in 2012, and estimate the difference in readmission and mortality rate responses for nonprofit hospitals with and without clinically trained executives. I find that a lack of clinical executives leads hospitals to improve quality more drastically after the incentive change. Further, I provide evidence that the more drastic response is due to non-clinical executives being more profit driven, and that clinical executives are not just a signal of underlying hospital objectives, but manage the hospital differently. 
		} 
	\end{abstract}

@@

    Various characteristics of firm leadership such as gender, age, and background have been shown to be correlated with firm decisions and performance. I investigate this topic in the context of hospital executives with differing backgrounds: some with training as a medical doctor and some without. Hiring physician executives can be controversial. On one hand, physicians bring a unique combination of clinical and administration expertise, and they have insights that can potentially improve hospital outcomes (\cite{Stajduhar_2023}, \cite{Ahmed_2022}). Alternatively, physicians may not always be adequately trained to enter into high level management positions, and therefore worsen hospital outcomes (\cite{HarvardBusinessReview2018}). In this paper, I investigate how having clinically trained executives in leadership affects hospital response to pay-for-performance incentives.

    I scrape executive team names and titles from publicly available Tax Form 990s, filled out by nonprofit firms in the US. Combining this with publicly available hospital characteristics, I construct a novel data set of nonprofit hospitals from 2010-2014 containing information on both executive team characteristics and quality metrics. Nonprofit behavior in the hospital industry directly affects health care consumers, as private nonprofit hospitals make up 50\% of all hospitals in the US, and staff, on average, 207 beds. For comparison, for-profits make up 36\% of hospitals and staff 107 beds (\cite{ASPE_2023}).

    The selection of leaders makes it difficult to interpret any direct comparison of differing executive teams as causal. Therefore, I leverage an exogenous shock to hospital incentives, pay-for-performance policies enacted in the US in 2012, and compare the change in readmission and mortality rates of executive teams with and without clinical executives. Under the assumption that the two types of teams would have had similar trends in readmission and mortality rates absent the incentive change, I find that a lack of clinically trained executives leads a hospital to respond more drastically to pay-for-performance incentives.

    Having found that hospital leadership does, in fact, matter to hospital behavior, I investigate several mechanisms that might contribute to this difference. First, \textcolor{red}{the more drastic decrease in readmissions by nonclinical teams is not driven by decreasing average patient complexity. There is no difference in patient case mix between clinical and nonclinical executive teams}. Further, I comparing each type of leadership team with for-profit hospitals reveals that the absence of clinical executives leads a nonprofit hospital to respond more similarly to a for-profit hospital. That is, having clinical training on a leadership team leads a hospital to act less profit-driven than their general training counterparts. Finally, I show that changes in a hospital's propensity to hire a clinical executive are not correlated with the pay-for-performance policies. Thus, I exploit such changes in leadership teams to decompose the effect of clinical executives into signaling (clinical executives signal underlying hospital preferences), or managing (clinical executives manage the hospital differently). I find that the entire difference in response is driven by a managing effect.

    Many researchers have documented a correlation between executive characteristics, specifically for CEOs, and firm performance, starting with \citeauthor{bertrand2003managing} (\citeyear{bertrand2003managing}), who show that manager fixed effects are an important driver of many firm behaviors and decisions. Female board members are associated with better oversight and more women executives (\cite{matsa2011chipping}; \cite{adams2009women}), but are negatively correlated with firm value (\cite{ahern2012changing}). Having more female executives is correlated with female employee wages and corporate strategies (\cite{flabbi2019female}; \cite{matsa2013female}); young male CEOs tend to be more aggressive in mergers and acquisitions, while those with military experience are less aggressive (\cite{levi2010deal}; \cite{benmelech2015military}); CEOs with general ability tend to receive higher pay and perform better (\cite{kaplan2012ceo}; \cite{custodio2013generalists}; \cite{adams2018director}; \cite{frydman2019rising}). Further, Chief Diversity Officers are not found to have any effect on hiring more diversely in universities (\cite{bradley2022impact}).

    Due to data limitations, many of these studies focus on large, publicly traded firms in the US. \citeauthor{brickley2010board} (\citeyear{brickley2010board}) has the only study to my knowledge answering this question in the US nonprofit context. Using exogenous variation in expected Medicare profits, the authors find that having an internal board of directors increases CEO compensation, and having physician board members decreases public donations (\cite{brickley2010board}). Additionally, two studies focus on hospital performance in other countries. \citeauthor{janke2019impact} (\citeyear{janke2019impact}) uses data from England to study whether CEOs affect hospital production, and find no association (\cite{janke2019impact}). However, \citeauthor{otero2022managers} (\citeyear{otero2022managers}) investigates the role of CEOs in public hospitals in Chile, and finds an 8\% decrease in mortality rates for hospitals with top managers. Further, a major contributor of this quality improvement is the exodus of older physicians as CEOs (\cite{otero2022managers}). I contribute to our knowledge of how executive team characteristics affect firm behaviors by (1) considering the important setting of nonprofit hospitals, which serve the majority of patients in the US, (2) considering a unique executive characteristic which is plausibly relevant, and (3) observing firm behavior in the presence of an exogenous incentive change.

    Additionally, I contribute to our understanding of how providers respond to pay-for-performance incentives in health care. The most in depth study of how hospitals respond to the Hospital Readmissions Reduction Program, one of the two large pay-for-performance initiatives, is \citeauthor{gupta2021impacts} (\citeyear{gupta2021impacts}). This study finds that hospitals decreased readmissions by 5\% and mortality rates by 2\% on average as a result of the program, confirming prior studies (\cite{mellor2017does}; \cite{ziedan2018essays}; \cite{ody2019decreases}; \cite{gupta2021impacts}). Around 40\% of the decrease in readmissions is due to selective patient practices. Further, market concentration and hospital systems play a large role in how hospitals respond to the program (\cite{kunz2024assessing}). Research on the other large pay-for-performance initiative, the Hospital Value Based Purchasing Program, generally finds that the program had no effect on underlying hospital quality (\cite{us2015hospital}; \cite{norton2018moneyball}; \cite{friedson2019so}). This paper contributes to our understanding of how characteristics of hospitals affect response to pay-for-performance incentives.

    \section{Setting and Data}

    \subsection{Nonprofit Hospital Executives}

    Nonprofit firms in the US differ from for-profits mainly in that for-profit firms distribute revenue to stakeholders, while nonprofits reinvest profit back into the firm. Hospitals are governed by a board of directors, whose role is to set broad goals and strategies, and provide general oversight. The board selects executives, the highest level of management of a firm, to carry out day-to-day operations. An executive team usually consists of at least a Chief Executive Officer (CEO) and Chief Financial Officer (CFO), but there is variation in how firms organize executive teams. Some hospital executives to specialize in health care settings by earning a degree in health care management, or an MBA specific to health care. There are executive positions that are often filled by someone with clinical training, such as a Chief Medical Officer (CMO), but medical doctors can pursue other top managerial positions as well such as CEO or president. While some doctors earn additional degrees before stepping into an executive role, this is not a necessary condition to becoming a physician executive.

    Hospital executive teams are understudied partly due to the inaccessibility of granular data. I construct a novel data set of not-for-profit hospitals in the US that contains names, titles, and positions of all board members and executives tied to the hospital from 2009-2015. I gather this data from publicly available Tax Form 990s, which every sufficiently large nonprofit files each year in the US. To my knowledge, this is the first large scale gathering of executive names from these forms.\footnote{\citeauthor{brickley2010board} (\citeyear{brickley2010board}) collects compensation data for a select number of hospitals.}

    The historical tax forms for all not-for-profits are housed by ProPublica.\footnote{https://projects.propublica.org/not-for-profits/} I use the NonProfit Explorer API to extract Employee Identification Numbers (EIN) of NFPs with a hospital National Taxonomy of Exempt Entities (NTEE) code. After filtering out associations and specialty hospitals, there are approximately 3,000 NFP EINs. Linking these EINs to other hospital information is crucial for the analysis. I therefore rely on matching by location and name of hospital in both the tax forms and AHA data. In the AHA data, I limit to general hospitals. Then, I confidently match 1200 EINs to an AHA identification number based on exact name matches within the same state. I assess the observable differences between matched and non-matched hospitals in Appendix \ref{app:matched}, and conclude that the samples are overall similar apart from an under-sampling of hospitals that belong to systems.

    Finally, I match each name with the National Plan and Provider Enumeration System (NPPES) database of all registered physicians. The goal is to match each physician with an NPI number in order to merge publicly available data on physician specialty, gender, and tenure. For any person who has a doctor title and uniquely matches the NPPES data, I record their NPI number. For anyone who has a doctor title and no matches, no doctor title and matches, or a doctor title and multiple matches, I use information on location and hospital name to manually match the NPI.

    For the purposes of this paper, I focus on executive teams as a whole instead of specific positions. The main reason for this is because the organization of teams varies so drastically across hospitals that a specific position, even if it exists in two different hospitals, can be meaningfully different. I drop anyone who is solely a board member, and focus on those who have executive in their title, or are labeled as a president or vice president of a department. I create various hospital-level characterizations based on their executive team, such as the existence of a clinically trained executive, the number of clinically trained executives, the number of total executives, and whether the hospital has a Chief Medical Officer. For my main sample, I limit to hospitals who do not change their propensity to hire a clinical executive for the duration of the sample. Summary statistics for these variables are shown in Table \ref{leader_sumstats}. On average, NFP hospitals in my sample have 4 executives and .4 clinically trained executives. Sixty-one percent of the sample never changes propensity to hire a clinically trained executive, and 33\% of NFPs in the sample have a Chief Medical Officer at some point.

    \subsection{Pay-for-Performance Policies Targeting Readmission and Mortality Rates}\label{sec:hrrp}

@@

    Therefore, I focus on outcomes of readmission and mortality rates, which are publicly available at the hospital-condition level in CMS Hospital Compare. I combine rates for pneumonia, AMI, and heart failure (the relevant HRRP conditions) into a weighted average based on the number of patients in each condition. Additionally, I include several relevant hospital characteristics from the AHA survey: number of beds, and whether the hospital is an academic medical center, is physician owned, or is system affiliated. Additionally, case mix index, a measure of overall patient complexity, comes from the CMS Case Mix Index files. Whether the hospital was penalized under HRRP or received payments under HVBP is found in the HCRIS data.

	\subsection{Summary Statistics}\label{sec:data}

    In Table \ref{tab:sumstats_samples}, I present a table of means of relevant variables for several hospital samples used throughout the analysis. First, there are 708 for-profit hospitals in the sample, shown in column (1). For-profit hospitals in the sample have 155 beds on average, 112-276 patients depending on the condition, and are highly likely to be affiliated with a system. Thirty six percent of for-profits were penalized for all three conditions at some point.

@@

    Columns 2-4 show variable means for not-for-profits in the sample, with all NFPs in column 2, NFPs that ever have a clinically trained executive in column 3, and NFPs that never have a clinically trained executive in column 4. NFPs are much less likely to be physician owned or system affiliated compared to FPs. They do, however, have similar bed size and case mix index to FPs. Twenty-two percent of all NFPs are penalized for all conditions. There are several differences among NFPs with and without clinically trained executives. Those with a clinically trained executive are larger in terms of beds and patients seen, are more likely to be an academic medical center, and see slightly more complex patients. Further, they are more likely to be penalized for every condition.

    Finally, columns 5-6 limit the sample of not-for-profits further using changes in executive teams. The NFPs included in this sample do not change their propensity to hire a clinically trained executive, they either always have the same number of clinically trained executives or always have no clinically trained executives. The differences in NFPs with and without clinically trained executives remains similar with this sample limitation. I discuss the implications of these differences for identification in Section \ref{sec:model}.

    I plot weighted average readmission and mortality rates across time for different hospital types in Figure \ref{fig:weighted_read_mort_graph}, where I adjust both rates by baseline patient complexity. Pay-for-performance incentive changes are shown as a dotted line between 2011 and 2012. For-profit hospitals have the highest readmission rates prior to 2012. While the readmission rate trend behaves similarly across hospital types prior to the programs, there are notable trend differences after the programs. For-profits decrease readmissions drastically. NFP trends depend on the presence of an executive MD, where those without a clinically trained executive respond drastically, more similarly to for-profits, than those with a clinically trained executive. In mortality rates, NFPs with a clinically trained executive have the lowest rates before the programs, but differentially continue with an increasing trend in mortality after the programs. While this is purely descriptive evidence, the trends in outcomes point to more similar behavior between for-profits and NFPs without a clinically trained executive in comparison to NFPs with a clinically trained executive.

@@

    \section{Effect of Executive Clinical Training on Hospital Response}\label{sec:clinical}

    My first goal is to estimate the effect of having clinical executive training on how hospitals change readmission and mortality rates in response to pay-for-performance incentives. Ultimately, I employ synthetic difference-in-differences, weighting more heavily non-clinical hospitals with readmission, mortality, and baseline case mix index that better match the clinical team trends. However, for clarity, I present the unweighted specification:

    \begin{equation}
    \label{eq:clinical}
    y_{ht} = \beta \text{ (clinical training x post 2012)}_{ht} + \gamma_{h} + \delta_t + \epsilon_{ht},
    \end{equation}

    \noindent where $y_{ht}$ is one of the outcome variables discussed in Section \ref{sec:hrrp}, and $\gamma_h$ and $\delta_t$ are hospital and time fixed effects, respectively.

     I rely on several assumptions to identify this effect. First, I assume that, absent the program enactments, hospitals with different leadership team characterization types would have behaved similarly in outcomes. While I see no evidence of pre-trends, Table \ref{tab:sumstats_samples} suggests that hospitals with a clinically trained executive are different than other hospitals along multiple characteristics. There may be concern that these differences between hospital types could be driving the results. Therefore, my main specification uses the synthetic difference-in-differences method established in \citeauthor{arkhangelsky2021synthetic} (\citeyear{arkhangelsky2021synthetic}). This method creates unit weights that more heavily factor comparison hospitals that are similar to the treat group hospitals prior to the programs. It also weighs time periods that balance pre-programs and post-programs for the comparison hospitals. Then, a simple two way fixed effects estimator is used with these weights, yielding a more robust average treatment effect that relies less on a strong parallel trends assumption (\cite{arkhangelsky2021synthetic}). I present the synthetic diff-in-diff estimates as the main results, and they are largely similar to the standard two way fixed effects results with no weights, shown in Appendix \ref{app:fullsample}. Further, the conceptual identification is the same as a standard difference-in-differences.

@@

    The estimated difference in readmission rate trends for all hospital types in response to the programs are shown in Figure \ref{fig:clinicalsynthdid}. In panel \ref{fig:read_synth_clinical}, I compare readmission rate responses among nonprofit hospitals with and without clinical training on their executive team. Estimate $\beta$ is represented in the difference in slope after 2012 when pay-for-performance incentives began. Nonprofits without a clinically trained executive lowered readmission rates by .24 ppts more than those with a clinically trained executive.

    I also directly estimate the difference in mortality rates among NFPs with and without clinical experience in panel \ref{fig:mort_synth_clinical}. Again, the difference is not statistically significant, but the coefficient compared to the true nulls in panels \ref{fig:mort_synth_clinical} and \ref{fig:mort_synth_plotc} indicates NFPs without a clinical executive potentially decrease mortality more aggressively than NFPs with a clinical executive. In Appendix \ref{app:condition}, I show that these results are mainly driven by pneumonia patients.
